\Title

### 2023-06-02 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

The present paper proposes a real-world application that extends a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}. Indeed, this previous model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures. This was a first demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation thanks to heterogeneous delays. Here, we extend these results to a more complex and natural setting. First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested. Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades. We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays. Using this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner. We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning. Indeed, once trained, the amount of event-driven computation could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task. In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.

>>>>

The present paper proposes a real-world application that extends a recent model of spiking neurons with heterogeneous synaptic delays~\cite{grimaldi_learning_2022}. Indeed, this previous model was trained to solve a simplified motion detection task on a synthetic event-based dataset generated by moving parameterized textures. This was a first demonstration that formal neurons can exploit the precise timing of spikes to improve neural computation thanks to heterogeneous delays. Here, we extend these results to a more complex and natural setting. First, we define the ecological cognitive task that the model must solve with the different datasets on which it will be tested. Instead of the synthetic textures that have been used to generate motion-driven stimuli, we use natural images and random movements that mimic biological saccades. We then develop the Heterogeneous Delays Spiking Neural Network (HD-SNN) model thanks to theoretically defined first principles and derive a learning rule to adapt heterogeneous synaptic delays. Using this detection task, we study the emergence of spatiotemporal spiking motifs when this single layer of spiking neurons is trained in a supervised manner. We investigate the efficiency of the motion detection mechanism and, in particular, its resilience to synaptic weight pruning. Indeed, once trained, the amount of event-driven computations could be drastically reduced by removing weak synapses while maintaining peak performance for the classification task. In this way, we will be able to show how such a model can provide an efficient solution to the energy/accuracy tradeoff.
